

---
title: "Data101 Project 2"
author: "Praveena Tallapragada"
date: "24/06/2019"
output:
  html_document:
    toc: yes
    toc_depth: '2'
---


#### Dataset - LA International Airport Passenger Traffic by Terminal

[Source File](https://catalog.data.gov/dataset/los-angeles-international-airport-passenger-traffic-by-terminal-756ee)

I have selected the "LA International Airport Passenger Traffic by terminal" dataset from Data.gov. This is a simple dataset with 6 variables and 5870 observations.
        I had selected this dataset for my Project 1 to see if I can know the busiest terminal with respect to the total number of passengers.
        For my second project, I would like to find any patterns in passenger traffic by gathering the monthly data for Tom Brady International Terminal.

This dataset has the following 6 columns:

1. DataExtractDate	

2. ReportPeriod	

3. Terminal	

4. Arrival_Departure	

5. Domestic_International	

6. Passenger_Count

#### EDA Principles

##### Formulate your question

I would like to find any patterns in passenger traffic by gathering the monthly data for the year 2018 for one domestic terminal and Tom Brady International Terminal.By observing the passenger traffic on a monthly basis, I hope to find a pattern.

##### Read in your Data

I will be using the following library's to load, view, clean and plot the data:

```{r setup,message = FALSE}

library(tidyverse)
library(dplyr)
library(ggplot2)

```

Load the source file from the working directory, get the working directory and ensure the source files is located in this path...

```{r, echo=T, results='hide'}

getwd()

```

Load the file in R...

```{r,echo=T, results='hide'}

LA_Airport_Source <- read_csv("Los_Angeles_International_Airport_-_Passenger_Traffic_By_Terminal.csv")

```

##### Check the packaging

```{r}
nrow(LA_Airport_Source)

ncol(LA_Airport_Source)
```

##### Run str()

```{r}

str(LA_Airport_Source)

```

##### Look at the top and the bottom of your data

```{r}
head(LA_Airport_Source[, c(2:6)],10)
tail(LA_Airport_Source[, c(2:6)],10) 

```

#### Cleaning and preparing data with dplyr

Source file contains data from January 2006 to March 2019. For every terminal, we have passenger counts for arrival or departure and whether it is domestic or international.

"DataExtractDate" and "ReportPeriod" are stored as characters instead of a date time variable. Also the time indicated in the variable 'Reportperiod' is always defaulted to '12:00:00 AM' and therefore it will not be of any use. I will have to remove it and maybe split the 'ReportPeriod' date into month and year to better understand the passenger traffic and find any patters if possible. I will remove 'DataExtractData' as I am not interested in knowing when the data was collected for this project.

To undertsand the data in the next three columns - "Terminal","Arrival_Departure" and "Domestic_International", I will get the unique list from these columns.

```{r}

unique(LA_Airport_Source$Terminal)

unique(LA_Airport_Source$Arrival_Departure)

unique(LA_Airport_Source$Domestic_International)

```


The three variables are categorical and hence it would be better to change the datatype to factor for all three of them.

Now lets check the range of passenger count, just to have an idea of number of passengers.


```{r}

range(LA_Airport_Source$Passenger_Count)

```

Check for NA's and complete cases

```{r}

which(is.na(LA_Airport_Source))

which(!complete.cases(LA_Airport_Source))

```

Remove "DataExtractDate" column

```{r}

LA_Airport_Modified <- LA_Airport_Source %>% select(ReportPeriod,Terminal,Arrival_Departure,Domestic_International,Passenger_Count) 

LA_Airport_Modified

```


Split the "ReportPeriod" column and remove the time.Remove data for the year 2019 as it is incomplete.

```{r}

LA_Airport_Modified$ReportPeriod <-  as.Date(LA_Airport_Modified$ReportPeriod, "%m/%d/%Y")

LA_Airport_Modified <- LA_Airport_Modified %>%
  mutate(Year = format(ReportPeriod, "%Y"),Month = format(ReportPeriod, "%m")) 

LA_Airport_Modified$Year <- as.numeric(LA_Airport_Modified$Year)

LA_Airport_Modified$Month <- as.numeric(LA_Airport_Modified$Month)

LA_Airport_Modified

LA_Airport_Mod <-  LA_Airport_Modified %>%
filter(Year !=  2019)

LA_Airport_Mod

str(LA_Airport_Mod)

```

Make the variables - "Terminal","Arrival_Departure" and "Domestic_International" as factors and try to spread them if needed.Replace space from Terminal names with underscroes.

```{r}

LA_Airport_Mod$Terminal <- gsub("\\s+","_",LA_Airport_Mod$Terminal)

LA_Airport_Mod$Terminal <- as.factor(LA_Airport_Mod$Terminal)

LA_Airport_Mod$Arrival_Departure <- as.factor(LA_Airport_Mod$Arrival_Departure)

LA_Airport_Mod$Domestic_International <- as.factor(LA_Airport_Mod$Domestic_International)

summary(LA_Airport_Mod)

```

Calculate the sum of passenger counts per terminal per year per month.

```{r}

Monthwise_Data <- LA_Airport_Mod %>%
  select(Terminal,Year,Month,Passenger_Count) %>%
  group_by(Terminal,Year,Month) %>%
  summarize(Passenger_sum = sum(Passenger_Count)) %>%
arrange(desc(Passenger_sum))

Monthwise_Data

D_I_perM <- LA_Airport_Mod %>%
  select(Domestic_International,Year,Month,Passenger_Count) %>%
  group_by(Domestic_International,Month) %>%
  summarize(Passenger_sum = sum(Passenger_Count)) %>%
  arrange(desc(Passenger_sum))

D_I_perM


```

#### ggplot2 - Plots

BoxPlot

```{r}

 boxplot(Monthwise_Data$Passenger_sum, col = "blue")

 boxplot(D_I_perM$Passenger_sum, col = "blue")

```
Histogram

```{r}

hist(Monthwise_Data$Passenger_sum, col = "red")

```
Scatterplot

```{r}


str(D_I_perM)

 with(D_I_perM, {plot(Month, Passenger_sum,col = as.factor(Domestic_International), xlab = "Month", ylab = "Passenger Count")
 legend = D_I_perM$Domestic_International})
 
 

```

ggplot - Tom Brady International Terminal 
```{r}


TB_Monthwise_Data <- LA_Airport_Mod %>%
  select(Terminal,Year,Month,Passenger_Count) %>%
  filter(Terminal == "Tom_Bradley_International_Terminal") %>%
  group_by(Year,Month) %>%
  summarize(Passenger_sum = sum(Passenger_Count)) %>%
  arrange(Year)

TB_Monthwise_Data

p1 <- ggplot(data = TB_Monthwise_Data) + 
  geom_line(mapping = aes(x = Month, y = Passenger_sum,  color = factor(Year)))+
  labs(x = "Month",y = "Total Passenger Count") +
  scale_x_continuous(breaks  = c(1,2,3,4,5,6,7,8,9,10,11,12))

p1

p2 <- ggplot(data = TB_Monthwise_Data, mapping = aes(x=factor(Month), y=Passenger_sum)) + 
  geom_boxplot() +
  geom_jitter(width = 0.25)

p2

p3 <- ggplot(data = TB_Monthwise_Data, mapping = aes(x = Month, y = Passenger_sum)) +
  geom_point() +
  geom_smooth(method = "lm", se=FALSE)

p3


p4 <- ggplot(LA_Airport_Modified, aes(ReportPeriod, Passenger_Count, colour = Arrival_Departure)) +
  geom_line()

p4

p5 <- ggplot(data = Monthwise_Data) + 
  geom_line(mapping = aes(x = Month, y = Passenger_sum,color = factor(Year))) + 
  facet_wrap(~ Terminal) +
  labs(x = "Month",y = "Total Passenger Count") +
  scale_x_continuous(breaks  = c(2,4,6,8,10,12)) 
  

p5

```

#### Summary

Using ggplot, I have tried to create 5 plots:

1. ggplot + geom_line: 

This visualizations displays Month on x axis and Passenger sum on y axis from TB_MOnthwise_data df. Year is displayed with different colors on the plot. As I mentioned in the beginning, I wanted to find patterns in passenger traffic by oberseving it on a monthly basis. We can clearly see that beginning 2006 to 2018, the passenger traffic decreases in february and steadily increases until July-August timeframe and then steadily decreases until november and slightly increases over december.Lowest passenger traffic is observed in february and highest passenger traffic is observed in July-August. Maybe during summer everyone is going on vacations and thats why we see the passenger traffic is highest in Tom Brady International Airport during the peak summer time.

2. ggplot + geom_boxplot + geom_jitter : 

This visualizations displays Month on x axis and Passenger sum on y axis from TB_MOnthwise_data df. I have tried to use the boxplot feature with jitter here to observe the Tom Brady International Airport data from 2006- 2018. the jitter feature helps to understand the spread and outliers much more conviniently. We can clearly make out that the median passenger trafic in summer is more than the median during the beginning or end of the year.

3. ggplot+ geom_point + geom_smooth:

I tried to plot the Tom Brady International Terminal data from 2006-2018 using the geom_point and geom_smooth just to try and understand what we can see and interpret.
I understand from this visualization that the passenger traffic increases from start to end of a year.

4. ggplot + geom_line:

In this visualization I have used the LA_Airport_Modified df.On the x axis, we have the Report Period and y axis we have the passenger count.The color display Arrival or departures.FRom the visualization, we can understand that the passenger counts have been high for arrivals than those of departures.


5. ggplot + geom_line + facet_wrap:

Using the facet wrap I have tried to visualize the passenger traffic from 2006-2018 for all the 11 terminals in LA International airport on a monthly basis.Tom Bradley international airport has seen a steady increase in passenger traffic over time.


